---
title: "TENxVisiumData"
author:
- name: Helena L. Crowell
  affiliation: Department of Molecular Life Sciences, University of Zurich, Zurich, Switzerland
output:
  BiocStyle::html_document:
    toc_float: true
package: TENxVisiumData
abstract: |
  The TENxVisiumData ExperimentHub package provides a collection of Visium spatial gene expression datasets by 10X Genomics. Data cover various organisms and tissues, and are formatted into objects of class SpatialExperiment.
vignette: |
  %\VignetteIndexEntry{TENxVisiumData}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r include = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, error = FALSE)
```

# Datasets

The `TENxVisiumData` package provides an R/Bioconductor resource for 
[Visium spatial gene expression datasets by 10x Genomics](https://support.10xgenomics.com/spatial-gene-expression/datasets), 
currently including ten samples from two organisms and seven tissues:

* [Human Breast Cancer (Block A Section 1)][HumanBreastCancer1]
* [Human Breast Cancer (Block A Section 2)][HumanBreastCancer2]
* [Human Heart][HumanHeart]
* [Human Lymph Node][HumanLymphNode]
* [Mouse Kidney Section (Coronal)][MouseKidney]
* [Mouse Brain Section (Coronal)][MouseBrainCoronal]
* [Mouse Brain Serial Section 1 (Sagittal-Anterior)][MouseBrainSagittalAnterior1]
* [Mouse Brain Serial Section 2 (Sagittal-Anterior)][MouseBrainSagittalAnterior2]
* [Mouse Brain Serial Section 1 (Sagittal-Posterior)][MouseBrainSagittalPosterior1] 
* [Mouse Brain Serial Section 2 (Sagittal-Posterior)][MouseBrainSagittalPosterior2]

These are provided as seven different datasets:

* HumanBreastCancer
* HumanHeart
* HumanLymphNode
* MouseKidneyCoronal
* MouseBrainCoronal
* MouseBrainSagittalPosterior
* MouseBrainSagittalAnterior

A list of currently available datasets can be obtained using the `ExperimentHub` interface:

```{r}
library(ExperimentHub)
eh <- ExperimentHub()
(q <- query(eh, "TENxVisium"))
```

# Representation

Each dataset is provided as a `r Biocpkg("SpatialExperiment")`, which extends the `r Biocpkg("SingleCellExperiment")` class with features specific to spatially resolved data.

Briefly, spatial coordinates are stored in the object's observation metadata (`colData`) and are accessible via the`spatialData()` accessor. Image-related data are stored inside the `int_metadata`'s `imgData` field and accessible with `imgData()`. For more details, see the package's vignette.

# Loading the data

```{r}
library(TENxVisiumData)
```

To retrieve a dataset, we can use the `TENxVisiumData()` function, which takes as input one of the dataset identifiers given above: 

```{r}
(spe <- TENxVisiumData(dataset = "HumanBreastCancer"))
```

Alternatively, data can loaded directly from Bioconductor's `r Biocpkg("ExerimentHub")` as follows. First, we initialize a hub instance and store the complete list of records in a variable `eh`:

```{r}
library(ExperimentHub)
eh <- ExperimentHub()
```

Using `query()`, we then identify any records made available by the `TENxVisiumData` package, as well as their accession IDs (EH1234):

```{r}
(q <- query(eh, "TENxVisium"))
```

Finally, we can load the data into R via `eh[[id]]`, where `id` corresponds to the data entry's identifier we'd like to load:

```{r}

```

<style type="text/css"> .smaller { font-size: 10px } </style>

# Session information {- .smaller}

```{r}
sessionInfo()
```

[HumanBreastCancer1]: https://support.10xgenomics.com/spatial-gene-expression/datasets/1.1.0/V1_Breast_Cancer_Block_A_Section_1
[HumanBreastCancer2]: https://support.10xgenomics.com/spatial-gene-expression/datasets/1.1.0/V1_Breast_Cancer_Block_A_Section_2
[HumanHeart]: https://support.10xgenomics.com/spatial-gene-expression/datasets/1.1.0/V1_Human_Heart
[HumanLymphNode]: https://support.10xgenomics.com/spatial-gene-expression/datasets/1.1.0/V1_Human_Lymph_Node
[MouseKidney]: https://support.10xgenomics.com/spatial-gene-expression/datasets/1.1.0/V1_Mouse_Kidney
[MouseBrainCoronal]: https://support.10xgenomics.com/spatial-gene-expression/datasets/1.1.0/V1_Adult_Mouse_Brain
[MouseBrainSagittalPosterior1]: https://support.10xgenomics.com/spatial-gene-expression/datasets/1.1.0/V1_Mouse_Brain_Sagittal_Posterior
[MouseBrainSagittalPosterior2]: https://support.10xgenomics.com/spatial-gene-expression/datasets/1.1.0/V1_Mouse_Brain_Sagittal_Posterior_Section_2
[MouseBrainSagittalAnterior1]: https://support.10xgenomics.com/spatial-gene-expression/datasets/1.1.0/V1_Mouse_Brain_Sagittal_Anterior
[MouseBrainSagittalAnterior2]: https://support.10xgenomics.com/spatial-gene-expression/datasets/1.1.0/V1_Mouse_Brain_Sagittal_Anterior_Section_2
